---
title: "Short demonstration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
---

# Example without location information

## Preparation

We use the `sample.ped` data from the `kinship2` package:
```{r}
data("sample.ped", package = "kinship2")
head(sample.ped)
```

And load helper packages:
```{r}
library(dplyr)
library(tibble)
```


Then we load the `popr` package:
```{r}
library(popr)
```

## Using `popr`

First, a population is build:
```{r}
pop <- with(sample.ped, load_individuals(pid = id, 
                                         is_male = (sex == 1), 
                                         pid_mom = mother, 
                                         pid_dad = father,
                                         progress = FALSE))
```

We check number of children:
```{r}
children_count <- as_data_frame(get_number_of_children(pop, progress = FALSE))
children_count %>% filter(boys_n > 0 | girls_n > 0)
```

We then build pedigrees:
```{r}
peds <- build_pedigrees(pop, progress = FALSE)
peds
```

```{r}
lapply(1L:pedigrees_count(peds), function(i) pedigree_size(peds[[i]]))
```

We can also get pedigree id's:
```{r}
ped_ids <- get_pedigree_id_from_pid(population = pop, pids = sample.ped$id)
table(ped_ids, sample.ped$ped)
```

We see that an individual is in its own pedigree as data does not suggest that it is linked to other individuals:
```{r}
peds[[3L]]
pid <- get_pids_in_pedigree(peds[[3L]])
sample.ped %>% filter(id == pid | father == pid | mother == pid)
```

Hence, be aware the the pedigrees constructed by `build_pedigrees` is (on purpose) only using the data provided.

We can inspect pedigrees (as induced by data):

```{r}
peds[[2L]]
pids <- get_pids_in_pedigree(peds[[2L]])
pids
sample.ped %>% filter(id %in% pids | mother %in% pids | father %in% pids)
children_count %>% filter(pid %in% pids)
```

We can also plot:
```{r}
plot(peds[[2L]])
g <- pedigree_as_igraph(peds[[2L]])
g
plot(g)
get_pedigree_as_graph(peds[[2L]])
```


# Example with location information

We add artificial [ETRS89](https://en.wikipedia.org/wiki/European_Terrestrial_Reference_System_1989) east and north coordinates (or a similar coordinate system). Remember, distances are Euclidean distances were ETRS89 east (`etrs89e`) and north (`etrs89n`) are regarded as $x$ and $y$ coordinates. Also, recall that the unit of the coordinates determine the unit of the distances.

```{r}
set.seed(1) # For reproducibility
sample.ped.coords <- as_data_frame(cbind(sample.ped, tibble(etrs89e = runif(nrow(sample.ped), 250, 650), etrs89n = runif(nrow(sample.ped), 650, 750))))
sample.ped.coords
```

First, a population is build:
```{r}
pop_coord <- with(sample.ped.coords, load_individuals(pid = id, 
                                                      is_male = (sex == 1), 
                                                      pid_mom = mother, 
                                                      pid_dad = father,
                                                      etrs89e = etrs89e,
                                                      etrs89n = etrs89n,
                                                      progress = FALSE))
```

We then build pedigrees:
```{r}
peds_coord <- build_pedigrees(pop_coord, progress = FALSE)
```

Now, prepare a population wide search tree (only for counting number of individuals):
```{r}
kdtree <- build_kdtree_from_population(pop_coord, max_leaf_size = 10)
```

Perform radius search of first individual:
```{r}
d <- meioses_distribution(population = pop_coord, search_tree_unrelated = kdtree,
                          max_leaf_size_pedigree = 10, 
                          pids = sample.ped.coords$id[1L], 
                          radii = c(100, 50, 100))
```

```{r}
d
```


 